---
title: "Updated ml_cov_search usage"
author: "mlcov"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Updated ml_cov_search usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
# Set to TRUE (or Sys.setenv(MLCOV_RUN_VIGNETTE = "true")) to run live searches
# and diagnostic plots. Default is FALSE so R CMD check stays fast.
run_live <- identical(tolower(Sys.getenv("MLCOV_RUN_VIGNETTE", "false")), "true")
```

# Purpose

This vignette is the shareable usage script for **mlcov 0.1.0**. It documents every new `ml_cov_search()` argument, shows how those arguments map onto the v2 manuscript methods, and (when `run_live` is `TRUE`) runs the packaged `tab33` example.

The manuscript simulation tables (Type I error, S1, S2 on 100 NONMEM replicates) **cannot** be reproduced from `tab33`. This file demonstrates that the package can invoke every Methods combination; it does not regenerate Figures 2–5.

To knit with live outputs:

```r
Sys.setenv(MLCOV_RUN_VIGNETTE = "true")
rmarkdown::render("vignettes/mlcov-updated-usage.Rmd")
```

```{r libs}
library(mlcov)
```

# Data

`tab33` is the packaged NONMEM-style example (one row per subject after `ID` is used). It is public demo data shipped with the package, not the NHANES simulation replicates from the paper.

```{r load-data}
data_file <- system.file("supplementary", "tab33", package = "mlcov")
data <- read.table(data_file, skip = 1, header = TRUE)

pop_param <- c("V1", "CL")
cov_continuous <- c(
  "AGE", "WT", "HT", "BMI", "ALB", "CRT", "FER", "CHOL",
  "WBC", "LYPCT", "RBC", "HGB", "HCT", "PLT"
)
cov_factors <- c("SEX", "RACE", "DIAB", "ALQ", "WACT", "SMQ")
```

# New arguments

`ml_cov_search()` now consolidates the eight specification variants
(learner × Lasso on/off) into one function.

```{r formals}
formals(ml_cov_search)
```

| Argument | Default | Role |
| --- | --- | --- |
| `use_lasso` | `TRUE` | Lasso pre-screen within each training fold |
| `lambda_lasso` | `"lambda.min"` | `"lambda.min"` or `"lambda.1se"` |
| `boruta_algorithm` | `"lightgbm"` | `"lightgbm"`, `"randomForest"`, `"xgboost"`, `"catboost"` |
| `boruta_pvalue` | `0.01` | Boruta significance level |
| `n_folds` | `5` | Outer 80/20 splits (caret) |
| `vote_threshold` | `2` | Keep covariates with `Freq >= vote_threshold` |
| `log_ebes` | `TRUE` | Log-transform EBEs (must be > 0) |
| `boruta_max_runs` | `200` | Boruta `maxRuns` |
| `seed` | `123` | RNG seed per parameter |

**Breaking changes from 0.0.2:** the default learner is LightGBM (was XGBoost); the default λ is `lambda.min` (was `lambda.1se`); SHAP/RMSE are no longer computed inside search.

# Manuscript methods vs package defaults

The v2 paper evaluates Boruta with random forest, XGBoost, LightGBM, and CatBoost, with and without Lasso, and with both λ rules. It recommends **Lasso `λ.min` + Boruta LightGBM**. Two settings differ from the implementation spec and are therefore explicit arguments:

```{r manuscript-notes, echo=FALSE}
notes <- data.frame(
  setting = c(
    "Learner",
    "Lasso",
    "Lambda",
    "Outer folds",
    "Vote rule",
    "Boruta p-value",
    "Boruta iterations",
    "LightGBM",
    "XGBoost",
    "CatBoost",
    "Random forest"
  ),
  manuscript = c(
    "Compared all four; LightGBM recommended",
    "With and without; Lasso recommended",
    "Both λ.min and λ.1se; λ.min recommended overall",
    "Five folds, 80/20 via caret",
    "More than two of five (majority)",
    "0.01 with Bonferroni (Boruta)",
    "Up to 100 iterations",
    "regression, RMSE, lr 0.1, 31 leaves, 200 rounds",
    "200 rounds, squared-error loss",
    "RMSE, depth 5, lr 0.05, 100 iterations",
    "Available; excluded from later analyses (Type I ≥ 40%)"
  ),
  package = c(
    "boruta_algorithm = \"lightgbm\" (default)",
    "use_lasso = TRUE (default)",
    "lambda_lasso = \"lambda.min\" (default)",
    "n_folds = 5 (default)",
    "vote_threshold = 2 (spec default); use 3 for the paper",
    "boruta_pvalue = 0.01 (default)",
    "boruta_max_runs = 200 (spec default); use 100 for the paper",
    "getImpLightGBM uses those hyperparameters",
    "getImpXgboost with nrounds = 200",
    "getImpCatBoost uses those hyperparameters",
    "boruta_algorithm = \"randomForest\" (not recommended)"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(notes)
```

Paper-faithful call (recommended workflow in the conclusion):

```{r manuscript-call, eval = FALSE}
result_paper <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  use_lasso = TRUE,
  lambda_lasso = "lambda.min",
  boruta_algorithm = "lightgbm",
  boruta_pvalue = 0.01,
  n_folds = 5,
  vote_threshold = 3,
  log_ebes = TRUE,
  boruta_max_runs = 100,
  seed = 123
)
```

# Worked examples

Live chunks below are gated by `run_live`. Calls are still shown so a colleague can copy them.

## Default search (LightGBM + Lasso λ.min)

```{r default-search, eval = run_live}
result <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors
)
print(result)
result$result_ML
result$result_5folds
result$settings
```

## Manuscript settings (majority vote, 100 Boruta runs)

```{r paper-search, eval = run_live}
result_paper <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  use_lasso = TRUE,
  lambda_lasso = "lambda.min",
  boruta_algorithm = "lightgbm",
  boruta_pvalue = 0.01,
  n_folds = 5,
  vote_threshold = 3,
  log_ebes = TRUE,
  boruta_max_runs = 100
)
print(result_paper)
```

## Contrast λ.1se and no Lasso

The paper found λ.1se more conservative for isolated categorical effects, and λ.min better when correlated true effects co-exist. `use_lasso = FALSE` is the unprocessed Boruta path.

```{r lambda-1se, eval = FALSE}
result_1se <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  boruta_algorithm = "lightgbm",
  use_lasso = TRUE,
  lambda_lasso = "lambda.1se"
)
```

```{r no-lasso, eval = FALSE}
result_nolasso <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  boruta_algorithm = "lightgbm",
  use_lasso = FALSE
)
```

## 0.0.2-compatible path (XGBoost + λ.1se)

```{r xgb-legacy, eval = FALSE}
result_xgb <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  boruta_algorithm = "xgboost",
  use_lasso = TRUE,
  lambda_lasso = "lambda.1se",
  vote_threshold = 2,
  boruta_max_runs = 200
)
```

## Other learners

Random forest is supported but not recommended (high Type I error in the paper). CatBoost runs only if the optional `catboost` package is installed.

```{r rf, eval = FALSE}
result_rf <- ml_cov_search(
  data,
  pop_param = pop_param,
  cov_continuous = cov_continuous,
  cov_factors = cov_factors,
  boruta_algorithm = "randomForest",
  use_lasso = TRUE,
  lambda_lasso = "lambda.min"
)
```

```{r catboost, eval = FALSE}
if (requireNamespace("catboost", quietly = TRUE)) {
  result_cat <- ml_cov_search(
    data,
    pop_param = pop_param,
    cov_continuous = cov_continuous,
    cov_factors = cov_factors,
    boruta_algorithm = "catboost",
    use_lasso = TRUE,
    lambda_lasso = "lambda.min"
  )
} else {
  message(
    "catboost is not installed. See ",
    "https://catboost.ai/en/docs/installation/r-installation"
  )
}
```

# Diagnostics (optional; not in the v2 paper)

SHAP and residual plots remain available as post-processing. They train an **XGBoost** interpreter on the selected covariates, regardless of the Boruta learner.

```{r shap, eval = run_live}
if (exists("result")) {
  shap_plots <- generate_shap_summary_plot(result, data, title.position = 0.5)
  shap_plots
}
```

```{r residuals, eval = run_live}
if (exists("result")) {
  generate_residuals_plot(data, result, pop_param = "CL")
  generate_residuals_plot(data, result, pop_param = "V1")
}
```

# Methods combinations checklist

Every combination in the manuscript Methods section is a single `ml_cov_search()` call:

```{r combo-grid, echo=FALSE}
algos <- c("lightgbm", "xgboost", "catboost", "randomForest")
grid <- expand.grid(
  boruta_algorithm = algos,
  use_lasso = c(TRUE, FALSE),
  lambda_lasso = c("lambda.min", "lambda.1se"),
  stringsAsFactors = FALSE
)
grid <- grid[grid$use_lasso | grid$lambda_lasso == "lambda.min", ]
grid$lambda_lasso[grid$use_lasso == FALSE] <- NA
rownames(grid) <- NULL
knitr::kable(
  grid,
  caption = "Call ml_cov_search() with each row. lambda_lasso is ignored when use_lasso is FALSE."
)
```

Copy-paste template:

```r
ml_cov_search(
  data, pop_param, cov_continuous, cov_factors,
  boruta_algorithm = "<algorithm>",
  use_lasso = TRUE,              # or FALSE
  lambda_lasso = "lambda.min",   # or "lambda.1se"
  n_folds = 5,
  vote_threshold = 3,            # manuscript majority; spec default is 2
  boruta_pvalue = 0.01,
  boruta_max_runs = 100,         # manuscript; spec default is 200
  log_ebes = TRUE,
  seed = 123
)
```
